Compositional equivalences based on Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2019-10-09 2016-Weak-FH-Bisim/WeakBisim.tex
commnets page 10, figure page 16

\subsection{Running Example}
\TODO{Rabea: trouver reference !!}
\TODO{refaire les commentaires}
To illustrate this work, we use the well-known Alternate bit protocol
(ABP) \TODO{Ref ?}, that provides safe transport of data between two processes,
over unsafe media. This protocol has been described and studied using
many approaches, because it is small enough, and can be easily
verified using exact finite models. Our version here is more
ambitious, encoding both the data and the control bits in a symbolic
way. One can imagine that this could easily be extended using bounded
integers, as in the sliding window protocol, or with more complex data
structures, in order to verify realistic protocols.

Fig. \ref{SimpleProt:Schema} shows the example principle: two unspecified
processes $P$ and $Q$ (Holes) communicate messages, with a data value
argument, through the two protocol entities $sender$ and
$receiver$. The underlying network is modelled by two media,
carrying the message together with a control bit on one way, and an
acknowledge back from $receiver$ to $sender$. The media can be perfect (safe), or
faulty (losing or duplicating messages) as shown in
Fig. \ref{schema:ABP-media}. Remark that the pLTSs shown in
Fig. \ref{schema:ABP-media} are in fact ``generic'' in the sense that
they will be instanciated in different ways depending on the number
and types of the message arguments. This is only a graphical syntax,
with an obvious semantics for building the corresponding pNets.

Fig. \ref{schema:ABP-composed} \TODO{To be commented}

@@

@@

\TODO{Eric: rewrite example with new Protocol}
\begin{example}\emph{An open-transition.}
  \label{OT:ABP}
At the upper level, the $ABP$ pNet has 2 holes and $Protocol$ has
a subnet. One of its possible open transitions (synchronizing the hole $P$
with the \emph{Protocol}) is:

 \smallskip\noindent
 $  OT_1  = \openrule{
%                       1 \xrightarrow{send(m,sb)}_{Sender} 2 ~~
%                            \xrightarrow{in(?m,?b)}_{M1}
      (P\mapsto send(m)); Pred=[m=m']; Post=\{s1\_m:=m\}
                      }
  %  {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}
    {\ostate{s_0,m_0,m_0,r_0} \OTarrow{\underline{send(m')}} \ostate{s_1,m_0,m_0,r_0}}
    $

@@

  The \emph{Protocol} pNet of Fig. \ref{schema:ABP-composed} has 4 controllers and no holes. One of its possible open-transition (the sender sending $(m,b)$ to the medium) is:

 \smallskip\noindent
 $  OT_2  = \openrule{
%                       1 \xrightarrow{send(m,sb)}_{Sender} 2 ~~
   %                           \xrightarrow{in(?m,?b)}_{M1}
   \emptyset; Pred=[send(SV2\_m, SV2\_b)=send(s_1\_m, s_1\_sb);in(SV2\_m, SV2\_b)=in(?m,?b)];  \\ Post=\{m_1\_m:=?m, m_1\_b:=?b, s_2\_sb:=s_1\_sb\}
                      }
  %  {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}
    {\ostate{s_1,m_0,m_0,r_0} \OTarrow{\underline{send(SV2\_m, SV2\_b)}} \ostate{s_2,m_1,m_0,r_0}}
    $

    \ERIC{Apres resolution de l'unification:}

     $  OT_2  = \openrule{
%                       1 \xrightarrow{send(m,sb)}_{Sender} 2 ~~
   %                           \xrightarrow{in(?m,?b)}_{M1}
      \emptyset; Pred=[s_1\_m=SV2\_m, SV2\_m=?m, s_1\_sb=SV2\_b, SV2\_b=?b];  \\ Post=\{m_1\_m:=?m, m_1\_b:=?b, s_2\_sb:=s_1\_sb\}
                      }
  %  {\ostate{00} \xrightarrow{\underline{\delta(x1)}} \ostate{10}}
    {\ostate{s_1,m_0,m_0,r_0} \OTarrow{\underline{send(s_1\_m, s_1\_b)}} \ostate{s_2,m_1,m_0,r_0}}
    $

@@

     $  OT_2  = \openrule{
      \emptyset; Pred=\True;  \\ Post=\{m_1\_m:=s_1\_m, m_1\_b:=s_1\_sb, s_2\_sb:=s_1\_sb\}
                      }
    {\ostate{s_1,m_0,m_0,r_0} \OTarrow{\underline{send(s_1\_m, s_1\_b)}} \ostate{s_2,m_1,m_0,r_0}}
    $

@@

\TODO{Add a reference to the construction (deduction tree) in the forte paper?}

@@

\TODO{Eric: full (???) result of the tool execution on the running example on the Simple Prot Spec, and maybe one example on the Impl. OTs presented in latex readable form.}

\subsection{Bisimulation for open pNets and Composability}
\label{section:bisimulation-PN}
As  our symbolic operational semantics provides an open automaton, we can apply the notion of
	strong (symbolic) bisimulation on automata to open pNets:
\begin{definition}[FH-bisimulation for open pNets]\label{def:bisim-pnets}
Two pNets are FH-bisimilar if there exist a relation between their associated 
automata that is an FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
\end{definition}
We can now prove that pNet composition  preserves
FH-bisimulation. More precisely, one can define two preservation
properties, namely 1) when one hole of a pNet is filled by two bisimilar other (open) pNets; and 2) when the same hole in two bisimilar pNets are
filled by the same pNet, in other words, composing a pNet with two
bisimilar contexts. The general case will be obtained by
transitivity of the bisimulation relation (Theorem~\ref{thm-equiv}).

>>>>

\subsection{Running Example}
\TODO{Rabea: trouver reference !!}
To illustrate this work, we use a simple transport protocol, that provides safe transport of data between two processes,
over unsafe media.

Fig. \ref{SimpleProt:Schema} shows the example principle: two unspecified
processes $P$ and $Q$ (Holes) communicate messages, with a data value
argument, through the two protocol entities $sender$ and
$receiver$. The underlying network is modelled by a medium entity transporting messages from the sender to the receiver, and that is able to detect some transport error, and signal it to the sender. The return $ack$ message from $receiver$ to $sender$ is supposed to be safe.

This is only a graphical syntax,
with an obvious semantics for building the corresponding pNets.

Fig. \ref{SimpleProt:Spec} shows the Schema of this system, and on the right side, its global specification, expressed as a pNet: it should be equivalent to a simple system where the holes would be connected by a perfect reliable medium.

@@

\TODO{Commenter la figure 2}

@@

\TODO{Eric: rewrite example with new Protocol}
\begin{example}\emph{An open-transition.}
  \label{OT:ABP}
At the upper level, the $SimpleSystem$ pNet has 2 holes and $SimpleProtocol$ as
a subnet. One of its possible open transitions (synchronizing the hole $P$
with the $Sender$ within the \emph{SimpleProtocol}) is:

 \smallskip\noindent
 $  OT_1  = \openrule{
      (P\mapsto p\_send(m)); Pred=[m=m']; Post=\{s1\_m:=m\}
                      }
    {\ostate{s_0,m_0,r_0} \OTarrow{\underline{in(m')}} \ostate{s_1,m_0,r_0}}
    $

@@

  The \emph{SimpleProtocol} pNet of Fig. \ref{SimpleProt:Spec} has 3 controllers and no holes. One of its possible open-transition (the sender sending $(m)$ to the medium) is:

 \smallskip\noindent
 $  OT_2  = \openrule{
   \emptyset; Pred=[send(SV2\_m)=send(s_1\_m);in(SV2\_m)=in(?m)];  \\ Post=\{m_1\_m:=?m\}
                      }
    {\ostate{s_1,m_0,r_0} \OTarrow{\underline{send(SV2\_m)}} \ostate{s_2,m_1,r_0}}
    $

    That can be simplified to:

@@

     $  OT_2  = \openrule{
      \emptyset; Pred=\True;  \\ Post=\{m_1\_m:=s_1\_m\}
                      }
    {\ostate{s_1,m_0,r_0} \OTarrow{\underline{send(s_1\_m)}} \ostate{s_2,m_1,r_0}}
    $

@@

In \cite{henrio:Forte2016}, we have shown a detailed example of the construction of a complex open transition, building a deduction tree using rules Tr1 and Tr2.

@@

\TODO{Eric: full (???) result of the tool execution on the running example on the Simple Prot Spec: Done, on the Counetr version !}

\TODO{:  and maybe one example on the Impl. OTs presented in latex readable form??}

\begin{figure}[t]
   \centerline{\includegraphics[width=9cm]{ATG/SPSpecOpen}}
   \caption{Specification OA (version with the error counter)}
   \label{SimpleProtCounter:SpecOA}
\end{figure}

\subsection{Bisimulation for open pNets and Composability}
\label{section:bisimulation-PN}
As  our symbolic operational semantics provides an open automaton, we can apply the notion of
	strong (symbolic) bisimulation on automata to open pNets:
\begin{definition}[FH-bisimulation for open pNets]\label{def:bisim-pnets}
Two pNets are FH-bisimilar if there exist a relation between their associated 
automata that is an FH-bisimulation and their initial states are in the relation, i.e. 
the predicate associated to the relation between the initial states is \True.
\end{definition}
\TODO{Predicate=True me pose un probleme... typiquement, on a besoin d'une ``condition initiale'', pour avoir des pNets data-oriented equivalents a leur version state-oriebted}

We can now prove that pNet composition  preserves
FH-bisimulation. More precisely, one can define two preservation
properties, namely 1) when one hole of a pNet is filled by two bisimilar other (open) pNets; and 2) when the same hole in two bisimilar pNets are
filled by the same pNet, in other words, composing a pNet with two
bisimilar contexts. The general case will be obtained by
transitivity of the bisimulation relation (Theorem~\ref{thm-equiv}).
